Physics-based protein structure prediction and design using the confinement method

### 2013-01-11 Paper/paper.tex
changed paper 11_6
Reviewed grammar and changed some ideas

Calculation of free energy differences is of central importance in the simulation of biochemical
systems. It is particularly difficult to calculate between pairs of macromolecular
conformations as well as a computationally expensive task with existing
methods. In this work, confinement approach is used to calculate absolute
free energies of biomolecular systems. This method provides two main advantages: it does not require a reaction coordinate or
transition path and it is fast to compute. Free energy calculated can be decomposed into a per
residue contribution in an 
approximate way. Per residue free energy allows us
to identify the reason behind conformational preferences in biomolecules. Through out the article we show its use in different
challenging modeling problems. In particular, we show its use in predicting the conformational
preference of chamaleon sequences (sequences with high sequence identity and different
folds). This sequence dependent conformational preferences and per residue free energy decomposition set the stage for the use of this 
method in protein design.

@@

In computational structural biology there are numerous cases where the free energy difference
between two well defined 
states plays an important role. Examples range from small to large conformational changes of proteins due to
ligand binding, change of pH or other conditions \cite{Meirovitch2007}, \cite{Chipot2007}, \cite{Jorgensen2004}. 
Free energy also plays an important role in the case of protein
folding. Ground breaking work of Christian B. Anfinsen and coworkers showed that the native
structures of small globular proteins have a unique, thermodynamically stable native structure with
their conformation at the global free energy minimum \cite{Anfinsen1973}. Regardless of the starting
point or even after unfolding 
it by changing different condition most proteins will go back to the native state after
reinstituting native like conditions. Often, the changes in free energy between different
conformations is small, but nonetheless crucial in favoring the native state
 since defects in protein folding may be
the molecular cause of a range of human genetic disorders. For example a misfolded protein known as
a
prion can misfold correctly folded proteins when entering a healthy organism.
\Alberto{I'm not sure I understand this here} \Arijit{Thus, Free energy can also act as a scale to identify between the misfolded and the native state of
the protein.} These facts further give free energy a special importance in structural biology.

However, the theoretical calculation of conformational free energy change becomes difficult if the
states of interest are very different \cite{Meirovitch2007}. In such scenario timescales involved in
such conformational transitions may
be beyond the sampling ability of classical molecular dynamics simulation.  Generally, in order to
calculate the free energy a pathway or reaction coordinate is needed with several intermediate structures.
Methods like umbrella sampling \cite{Torrie1977}, thermodynamic integration \cite{Tironi1994} along with classical MD can suffer from
overlap problem and requires a larger computational effort. The idea of a reaction coordinate became
even more
complicated in the case of protein folding.  As free energy landscape for protein folding 
is rough \cite{Dill1997}, \cite{Dill2008}, the molecule 
can get trapped in an energy well during conformational sampling. Even with 
advanced methods like replica exchange molecular dynamics, the three dimensional structure may be
trapped in a local minima for a considerable time, giving rise to nonconverged distributions where
misfolded states might be selected as native like
% (give the example of human pin1ww domain from Benout Roux and Klaus Schultan).

Calculation of free energy have been successfully attempted by a number of groups \cite{Meirovitch2007}, 
\cite{Ytreberg2006} - \cite{Zheng2008}. In recent years methods like the Reference system method \cite{Ytreberg2006}, 
Decativated Morphing \cite{Park2008}, Orthogonal Space Random Work \cite{Zheng2008} or the Confinement 
Method \cite{Tyka2006}, \cite{Cecchini2009} to name a few have been used to calculate free energy
differences.
In this work we have applied the confinement method which was originally devoled by Tyka et. al. \cite{Tyka2006} and
Cecchini et. al. \cite{Cecchini2009}. This method relies on the fact that the free energy difference is a state function and 
thus it is independent of the sampling path. This approach uses a thermodynamic cycle where first, non-harmonic degrees of
freedom are removed by applying a series of restraints to the system. Then the free energy of the
remaining harmonic system is obtained using a normal mode calculation and combined with the
non-harmonic part to give the total absolute free energy.  Previously, this method was applied to
calculate the conformational free energy of a 16 amino acid residue peptide, known as BHP.  We have
first reproduced that known result. Next we have applied this method to larger proteins for the
first time. The free energy difference of pairs of conformations in proteins with similar sequence but completely
different fold was calculated. \Alberto{i made this part too long, have to summarize here and define
better inside text?-->} Encouraged by the result, we applied the method to help tackle the
problem of protein conformation ranking. We did this by selecting different targets from the
Critical Assessment of Structure Prediction (CASP) event. In this event, different research groups
try to find the 3D structure of a protein given a sequence in a blind way using their methodologies. Upon comparing
with the real structure it is often seen that several groups are able to generate better structures
than the ones they submit, but are not able to identify them. This is known as the ranking problem,
and this method is able to correctly rank structures in the majority of cases we have tried.
 Finally, we have been exploring the ability of working on a per residue basis to identify which
 residues are the major players in the stabilization/destabilization of a pair of conformations. Our
 initial experience makes us thinkg that this method might be helpful in identifying key residues
 for protein design.

@@

The results of these calculations are presented in Figure~\ref{fig:orban}. Confinement method
identifies the correct structure for all five sequences. One of the hypothesis for fold switching
of chamelon sequence is that the structural transitions require states with diminished stability.
It is widely believed that if the free energy of the native state and the alternative state is within
 a range of $5 Kcal/Mol$ then it can quickly change fold when the stability of the native state 
decreases. The stability of the native state can decrease for a number reason ranging from chemical 
modification, breaking of disulphide bonds or mutations as in this case. The free energy differences, that
came out from our calculation ranges from
around 3.5 to 5.0 kcal/mol, which is consistent with the above hypothesis \cite{He2008}, \cite{{Alexander2009}}.

@@

\begin{table}
\begin{center}
\caption{Analysis of per residue free energy decomposition of GA95 and GB95 sequence reveal that certain residues prefer
either $3 \alpha$  or $4 \beta + \alpha$ structure. These observations and the reason behind such conformational preferences are listed in this
table.}
\label{table:orban_perresidue}
\begin{tabular}{| l | l | l |}
\hline
    Information from per                                 & What is happening in           & What is happening in                      \\ 
    residue plot                                         & $3 \alpha$ conformation        & $4 \beta + \alpha$  conformation          \\ \hline
    Leu-7 favors                                         & hydrophobic residue is         & oriented towards the                      \\ 
    $4 \beta + \alpha$ (Figure~\ref{fig:orban_full}(A))  & exposed to the solvent         & hydrophobic core                          \\ \hline
    Gln-11 stabilizes $3 \alpha$                   & Gln-11 has H-bond                         & no such H-bond                           \\ 
    (Figure~\ref{fig:orban_full}(B))           &  with Glu-15                              &                                           \\ \hline 
    Glu-14 favors $4 \beta + \alpha$           & hydrophilic residue partially  & completely exposed                                    \\       
    (Figure~\ref{fig:orban_full}(E))           & exposed to the solvent         & to the solvent                                        \\ \hline        
    Ala-26 favors  $4 \beta + \alpha$          & hydrophobic residue            & oriented towards the                                   \\ 
     (Figure~\ref{fig:orban_full}(C))          & oriented towards surface       & hydrophobic core                                       \\ \hline
    Ile-49 favors $3 \alpha$                   & hydrophobic interaction        & sidechain exposed                                       \\
    (Figure~\ref{fig:orban_full}(H) and (I))   & inside the protein             & to the solvent                                          \\ \hline 
    Ile-49 favors GA more                      & more hydrophobic               &  unchanged                                              \\
    than GB                                    & interactions with residues     &                                                         \\ 
    (Figure~\ref{fig:orban_full}(H) and (I))  &  at 20, 30 in GA                &                                                         \\ \hline
    Val-54 favors $4 \beta + \alpha$           & hydrophobic residue            & hydrophobic interaction                                \\
    Figure~\ref{fig:orban_full}(D))             & exposed to the solvent        & inside the protein                                      \\ \hline
    Glu-56 favors $3 \alpha$                   & hydrophilic residue partially  & completely exposed                                    \\
    (Figure~\ref{fig:orban_full}(G))            & exposed to the solvent         & to the solvent                                        \\ \hline
    Phe-30 in GB sequence                      & exposed to the solvent         & oriented towards the                                  \\
    stabilizes $4 \beta + \alpha$                  & in protein surface             & hydrophobic core                                      \\ \hline   
    Tyr-45 in GB favors                      & No H-bonding                   & H-bond with Asp-47                                    \\  
    $4 \beta + \alpha$ (Figure~\ref{fig:orban_full}(K))&                 &                                                       \\ \hline
    Asp-47 in GA sequence                      & H-bond with Lys-50             & No H-bond, No Tyr-45                                  \\  
    favors $3 \alpha$                          &                         & in GA sequence                                         \\ \hline
\end{tabular}
\end{center}
\end{table}

\begin{figure}
\includegraphics[width=3.5 in,height=4 in]{orban.pdf}
\label{fig:orban}
\caption{Confinement method correctly predicts the structural preferences of six chameleon
sequences. (A) The six sequences used in this study. (B) Each sequence adopts either a Protein
G-like fold (denoted GB) or a three-helix bundle fold (denoted GA). The relative free energies of
the two folds are reported for each sequence.}
\end{figure}

\begin{figure}
\includegraphics[width=6.4 in,height=5.3 in]{orban_full.pdf}
\label{fig:orban_full}
\caption{Sidechain orientation of some important residues which stabilize/destabilize either of the $4 \beta + \alpha$ or $3 \alpha$ conformation.
The details are in Table~\ref{table:orban_perresidue} and the main text.  
A. Leu-7.
B. Gln-11 has H-bond with Glu-15 in $3 \alpha$, which is absent in $4 \beta + \alpha$
C. Ala-26
D. Val-54.
E. Glu-14.
F. plot of free energy difference for GA95 and GB95 sequence. Positive peak stabilize $3 \alpha$ form and negative peak 
stabilize $4 \beta + \alpha$. 
G. Glu-56 .
H. sidechain orientation of Hydrophic Ile-49, along with Leu-20 and Ile-30 with GA95 sequence
I. sidechain orientation of Hydrophic Ile-49, along with Ala-20 and Phe-30 with GB95 sequence
J. Tyr-45 forms H-bond with Asp-47 with GB95 sequence and $4 \beta + \alpha$ conformation
K. No H-bond possible between Leu-45 and Asp-47 with GB95 sequence and $4 \beta + \alpha$ conformation .}  
\end{figure}

%\begin{table}
%\caption{Confinement method correctly reproduces the structural preferences of a series of
%designed peptides. Each sequence adopts either a Protein G-like fold ($4 \beta + \alpha$) or forms a
%helix bundle ($3 \alpha$).}
%\label{table:orban}
%\begin{center}
%\begin{tabular}{l l l}\hline
%Changes in & $\Delta G$ (kcal/mol)    & $\Delta G$ (kcal/mol)     \\
 %Sequences & ($3 \alpha$ fold)        & ($4 \beta + \alpha$ fold) \\ \hline
%GA95       & $0.000$                  & $2.901 \pm 0.475$         \\
%GB95       & $4.365 \pm 0.460$        & $0.000$                   \\
%GA88       & $0.000$                  & $ 3.941 \pm 0.517$        \\
%GB88       & $5.576 \pm 0.464$        & $0.000$                   \\
%GA98       &  &                           \\ \hline
%GB98       &                          &                           \\ \hline
%\end{tabular}
%\end{center}
%\end{table}

@@

This method is particularly well suited to pick out the native/native like structure from misfolded or decoys. We have
tested this using different models from CASP (critical assessment of structure
prediction). CASP is a blind test in which different groups world wide apply their
methods to predict the 3D structure of proteins given their sequence. This is done with a strict 3
week limit on each target (3 day for servers) and each group is allowed to submit 5 possible
structures (ranked from better to worst). In our role as assessors during the CASP8 and CASP9 \cite{MacCallum2011}
competition we observed no correlation at all between the ordering of structures submited by the
groups and the real ranking compared to an experimental model \cite{Kryshtafovych2011}. The consequences of this go beyond
those five structures; the deeper meaning is that groups producing ensembles of structures are
generating structures that are better than the ones they submit, but they do not know about it. In
fact, when querrying different groups after the results were known, most groups agree that this is
the case. Beyond the CASP problem, this reflects on the modelers ability to correctly rank order
models in many different environments, from structures for drug design leads to designing more
stable proteins or peptide mimetics such as peptoids. One of the main culprits of this lack of
accuracy is the fact that ranking is often done via a potential energy function, which in many cases
lacks an entropy component. Other initiatives including knowledge based potentials do use some sort
of free energy to rank order, but its accuracy is not enough. Our method provides a physics based
solution to this problem. We have tried two experiments centered on CASP. First we tried to rank
order some structures from the previous CASP9 experiment. We have tested different options for that purpose.
First, we rank order submition from same group. Then we rank submition from different groups. Our second test was 
to check whether this method can do better quality assessment/metaprediction than the best group in that catagory in CASP9. 
In both methods, our ranking is determined as a free energy measure,
and it is compared with the ranking given by a geometrical comparison between native and submitted
models called GDT\_TS (Global distance test score \cite{Zemla2003}). 
GDT\_TS represents the average percentage of residues that are in close proximity in two
structures optimally superimposed using four different distance cutoffs (1, 2, 4 and 8 Å).

We have chosen the inital models on which to test the methods based on server groups that have
traditionally done good in past CASP events. There is a high correlation between the structures that
are chosen for this method and its predictive accuracy. In particular, when GDT\_TS scores are below
50, it is difficult for us to say anything about the models. Our approach is simple, with every
subset of structures we choose, we use the confinement method in a pairwise procedure, and then rank
the structures. In all cases, we have calculated the free energy with respect to the native structure either obtained from NMR or 
x-ray crystallography but we do not need to make comparisons to native in order to do rank ordering.

\subsubsection{confinement method can rank different models submitted by the same group and distinguish the native structure}

The question that we want to answer here is whether we can rank models submitted by the same group.In all cases, we have calculated the free energy
difference between the native structure and the submitted models as well. 
%Identifying the native state from a set of models is a very easy test, that most metapredictors
%already handle correctly, but it is also a first proof of principle. 
First we choose a protein BVU3908 from Bacteroides vulgatus whose PDB id and CASP target code are
2L01 and T0559 respectively. The native structure of this 69 residue protein was solved using NMR.
The best predictor group for this target was "BAKER-ROSETTASERVER".
We initlally checked similarity between the models submitted by "BAKER-ROSETTASERVER" and discarded two of them from the 
analysis on the basis of being too similar to some of the other models. The GDT\_TS score and rmsd
values as shown in Figure~\ref{fig:T0559} indicate that the model 1 was predicted correctly, 
whereas the order of model 3 and 5 was wrong. Main difference
between the model 3 and the rest of the model is that the orientation of first alpha helix of model 3 is opposite.
On the other hand, as shown in Figure~\ref{fig:T0559} confine and release method not only can differentiate the native 
structure from the submitted models , it can also correctly rank all the models like their The GDT\_TS score.

To further test the method we have taken the example of protein BT2368 from Bacteroides thetaiotaomicron. The PDB id of this 
74 amino acid residue protein is 2L02 and CASP target code is T0560. We have compared the free energy difference between 
the native structure and the two of the five submitted models from the group "Splicer". The remaining three models are discarded as 
again they are similar to the rest of the models. The comparison of GDT\_TS score as shown in 
Figure~\ref{fig:T0560} and published in final CASP result is only for models with residues 3 to 66. In order to keep
consistency, we have also done our analysis with models and crystal structure consisting residues 3 to 66. As expected,
the native state is identified correctly, and the two other models are ranked in the correct order in accordance with the GDT\_TS score.

\subsubsection{Confinement method can rank models submitted by different groups}

Our next test was between models from the best submissions of different groups. 
For this purpose we have chosen the x-ray crystallographic structure of 
fas apoptosis inhibitory protein molecule whose pdb id and CASP target codes are 3MX7 and T0540 respectively.
This protein contain 90 amino acid residues with 8 beta strands. We have chosen best models from groups
"LTB" and "Mufold" for our analysis, which are labelled as Model 1 and Model 2 in Figure~\ref{fig:T0540}.
As presented in  Figure~\ref{fig:T0540} we found that this method is able to correctly order the models which matches
with the GDT\_TS score.

\subsubsection{Per residue free energy calculation identify the residues that are responsible for difference between two conformation}

In order to understand the per residue contribution towards the free energy, we have first calculated the total free energy of NMR 
structure of a domain of adhesion exoprotein from Pediococcus pentosaceus and its submitted CASP9 model. The 
pdb id and CASP code of this 79 residue protein are 2KWY and T0569 respectively. We have compared the native structure with the
best model structure submitted for this target in CASP9. This model structure was submitted by the group " Mufold". All the
five model structures submitted by this group was found to be close, within a GDT\_TS score  of 72 to 78. Thus we proceed by 
comparing only the native structure and
the best predicted model of this target. Visualization upon superimposition of two structures reveals that
the difference between the two structure is at the region consist of residues 48 to 65, where model1 is much more disordered compare to the
native structure(see Figure~\ref{fig:T0569}). Other regions look similar for both the cases with of model1 is within 2.6 \AA backbone
rmsd value of the native structure. We found that, confinement method can still differentiate between the native structure and 
the generated model. We have also calculated the per residue free energy with a aim that this can identify the residue which are responsible
for such free enerfy difference. The per residue free energy decomposition is shown in  Figure~\ref{fig:T0569_per_residue}(A). We indeed 
found that two hydrophobic residues Val-59 and Ile-61 are oriented towards the solvent in the protein 
surface in the generated model (Figure~\ref{fig:T0569_per_residue}(B). This the region, where there is a beta sheet in the original crystal structure, which is absent in the 
model. There are also some region, which stabilizes the generated model compare to that of the crystal. A close inspection reveal that 
the hydrophilic Glu in position 67 is completely exposed to the protein surface and solvent in the best model, whereas it is partially exposed in the 
crystal. This is presented in Figure~\ref{fig:T0569_per_residue}(C). The peak at position 76 is due to the H-bond between Lys-76 and Asp-11, 
which is absent in original crystalographic structure (Figure~\ref{fig:T0569_per_residue}(D)).

\subsubsection{Confinement method fails ?}

>>>>

The calculation of free energy differences is of central importance in the simulation of biochemical
systems. It is particularly difficult to calculate between pairs of macromolecular
conformations as well as a computationally expensive task with existing
methods. In this work, the confinement approach is used to calculate absolute
free energies of biomolecular systems. The method provides two main advantages: it does not require a reaction coordinate or
transition path and it is fast to compute. The free energy calculated can be decomposed into a per
residue contribution in an 
approximate way. Per residue free energy allows us
to identify the reason behind conformational preferences in biomolecules. Through out the article we show its use in different
challenging modeling problems. In particular, we show its use in predicting the conformational
preference of chamaleon sequences (sequences with high sequence identity and different
folds). This sequence dependent conformational preferences and per residue free energy decomposition set the stage for the use of this 
method in protein design.

@@

In computational structural biology there are numerous cases where the free energy difference
between two well defined 
states plays an important role. Examples range from small to large conformational changes of proteins due to
ligand binding, change of pH or other conditions \cite{Meirovitch2007}, \cite{Chipot2007}, \cite{Jorgensen2004}. 
Free energy also plays an important role in the case of protein
folding. The ground breaking work of Christian B. Anfinsen and coworkers showed that the native
structures of small globular proteins have a unique, thermodynamically stable native structure with
their conformation at the global free energy minimum \cite{Anfinsen1973}. Regardless of the starting
point or even after unfolding 
it by changing different condition most proteins will go back to the native state after
reinstituting native like conditions. Often, the changes in free energy between different
conformations is small, but nonetheless crucial in favoring the native state
 since defects in protein folding may be
the molecular cause of a range of human genetic disorders. For example a misfolded protein known as
a
prion can misfold correctly folded proteins when entering a healthy organism.
\Alberto{I'm not sure I understand this here} \Arijit{Thus, Free energy can also act as a scale to identify between the misfolded and the native state of
the protein.} These facts further give free energy a special importance in structural biology.

However, the theoretical calculation of conformational free energy change becomes difficult if the
states of interest are very different \cite{Meirovitch2007}. In such scenario timescales involved in
such conformational transitions may
be beyond the sampling ability of classical molecular dynamics simulation.  Generally, in order to
calculate the free energy a pathway or reaction coordinate is needed with several intermediate structures.
Methods like umbrella sampling \cite{Torrie1977}, thermodynamic integration \cite{Tironi1994} along with classical MD can suffer from
overlap problem and require a large computational effort. The idea of a reaction coordinate became
even more
complicated in the case of protein folding.  As the free energy landscape for protein folding 
is rough \cite{Dill1997}, \cite{Dill2008}, the molecule 
can get trapped in an energy well during conformational sampling. Even with 
advanced methods like replica exchange molecular dynamics, the three dimensional structure may be
trapped in a local minima for a considerable time, giving rise to nonconverged distributions where
misfolded states might be selected as native like
% (give the example of human pin1ww domain from Benout Roux and Klaus Schultan).

The calculation of free energy has been successfully attempted by a number of groups \cite{Meirovitch2007}, 
\cite{Ytreberg2006} - \cite{Zheng2008}. In recent years methods like the Reference system method \cite{Ytreberg2006}, 
Decativated Morphing \cite{Park2008}, Orthogonal Space Random Work \cite{Zheng2008} or the Confinement 
Method \cite{Tyka2006}, \cite{Cecchini2009} to name a few have been used to calculate free energy
differences.
In this work we have applied the confinement method which was originally devoled by Tyka et. al. \cite{Tyka2006} and
Cecchini et. al. \cite{Cecchini2009}. This method relies on the fact that the free energy difference is a state function and 
thus it is independent of the sampling path. This approach uses a thermodynamic cycle where first, non-harmonic degrees of
freedom are removed by applying a series of restraints to the system. Then the free energy of the
remaining harmonic system is obtained using a normal mode calculation and combined with the
non-harmonic part to give the total absolute free energy.  Previously, this method was applied to
calculate the conformational free energy of a 16 amino acid residue peptide, known as BHP.  We have
first reproduced that known result. Then we have applied this method to larger proteins for the
first time. The free energy difference of pairs of conformations in proteins with similar sequence but completely
different fold was calculated. \Alberto{i made this part too long, have to summarize here and define
better inside text?-->} Encouraged by the result, we applied the method to help tackle the
problem of protein conformation ranking. We did this by selecting different targets from the
Critical Assessment of Structure Prediction (CASP) event. In this event, different research groups
try to find the 3D structure of a protein given a sequence in a blind way using their methodologies. Upon comparing
with the real structure it is often seen that several groups are able to generate better structures
than the ones they submit, but are not able to identify them. This is known as the ranking problem,
and this method is able to correctly rank structures in the majority of cases we have tried.
 Finally, we have been exploring the ability of working on a per residue basis to identify which
 residues are the major players in the stabilization/destabilization of a pair of conformations. Our
 initial experience makes us thinkg that this method might be helpful in identifying key residues
 for protein design.

@@

The results of these calculations are presented in Figure~\ref{fig:orban}. The confinement method
identifies the correct structure for all five sequences. One of the hypothesis for fold switching
of chamelon sequence is that the structural transitions require states with diminished stability.
It is widely believed that if the free energy of the native state and the alternative state is within
 a range of $-5 Kcal/Mol$ then  it can quickly change fold when the stability of the native state 
decreases. The stability of the native state can decrease for a number reason ranging from chemical 
modification, breaking of disulphide bonds or mutations as in this case. The free energy differences, that
came out from our calculation ranges from
around 2.9 to 5.7 kcal/mol, which is consistent with the above hypothesis \cite{He2008}, \cite{{Alexander2009}}.

@@

\begin{table}
\begin{center}
\caption{Analysis of per residue free energy decomposition of GA95 and GB95 sequence reveal that certain residues prefer
either $3 \alpha$  or $4 \beta + \alpha$ structure. These observations and the reason behind such conformational preferences are listed in this
table.}
\label{table:orban_perresidue}
\begin{tabular}{| l | l | l |}
\hline
    Observation                                & $3 \alpha$ structure           &    $4 \beta + \alpha$ structure                       \\ \hline
    Leu-5 and Leu-7 favor                      & hydrophobic residues are       & oriented towards the                                  \\ 
    $4 \beta + \alpha$                         & exposed to the solvent         & hydrophobic core                                      \\ \hline
    Gln-11 favors $3 \alpha$                   & Gln-11 has H-bond with Glu-15  & no such H-bond                                        \\ \hline
    Glu-14 favors $4 \beta + \alpha$           & hydrophilic residue partially  & completely exposed                                    \\       
                                               & exposed to the solvent         & to the solvent                                        \\ \hline        
    Ala-26 favors  $4 \beta + \alpha$          & hydrophobic residue oriented   & oriented towards the                                   \\ 
                                               & towards surface                & hydrophobic core                                       \\ \hline
    Ile-49 favors $3 \alpha$                   & hydrophobic interaction        & exposed to the                                          \\
                                               & inside the protein             & solvent                                                \\ \hline 
    Val-54 favors $4 \beta + \alpha$           & hydrophobic residue exposed    & hydrophobic interaction                                \\
                                               & to the solvent                 & inside the protein                                      \\ \hline
    Glu-56 favors $3 \alpha$                   & hydrophilic residue partially  & completely exposed                                    \\
                                               & exposed to the solvent         & to the solvent                                        \\ \hline
    Phe-30 in GB sequence                      & exposed to the solvent         & oriented towards the                                  \\
    favors $4 \beta + \alpha$                  & in protein surface             & hydrophobic core                                      \\ \hline   
    Tyr-45 in GB Sequence                      & No H-bonding                   & H-bond with Asp-47                                    \\  
    favors $4 \beta + \alpha$                  &                                &                                                       \\ \hline
    Asp-47 in GA sequence                      & H-bond with Lys-50             & No H-bond, No Tyr-45                                  \\  
    favors $3 \alpha$                                 &                         & in GA sequence                                         \\ \hline
\end{tabular}
\end{center}
\end{table}

\begin{figure}
\includegraphics[width=3.5 in,height=4 in]{orban.pdf}
\label{fig:orban}
\caption{The confinement method correctly predicts the structural preferences of six chameleon
sequences. (A) The six sequences used in this study. (B) Each sequence adopts either a Protein
G-like fold (denoted GB) or a three-helix bundle fold (denoted GA). The relative free energies of
the two folds are reported for each sequence.}
\end{figure}

\begin{figure}
\includegraphics[width=6.5 in,height=5.4 in]{orban_full.pdf}
\label{fig:orban_full}
\caption{Orientation of residues of different sidechains in $4 \beta + \alpha$ and $3 \alpha$ form
A. Orientation of Hydrophobic residues Leu-5 and Leu-7.
B. Gln-11 has H-bond with Glu-15 in $3 \alpha$, which is absent in $4 \beta + \alpha$
C. Hydrophobic Ala-26
D. Hydrophobic Val-54.
E. Hydrophilic Glu-14.
F. The plot of free energy difference for GA95 and GB95 sequence. Positive peak stabilize $3 \alpha$ form and negative peak 
stabilize $4 \beta + \alpha$. 
G. Hydrophilic Glu-56 .
H. Orientation of Hydrophic Ile-49.
I. With GA95 sequence, Ile-49 have hydrophic interaction with both Leu-20 and Ile-30.
J. Larger Phe-30 in GB sequence, move towards the protein surface in $4 \beta + \alpha$.
K. Tyr-45 form H-bond with Asp-47 in $4 \beta + \alpha$.}  
\end{figure}

%\begin{table}
%\caption{The confinement method correctly reproduces the structural preferences of a series of
%designed peptides. Each sequence adopts either a Protein G-like fold ($4 \beta + \alpha$) or forms a
%helix bundle ($3 \alpha$).}
%\label{table:orban}
%\begin{center}
%\begin{tabular}{l l l}\hline
%Changes in & $\Delta G$ (kcal/mol)    & $\Delta G$ (kcal/mol)     \\
 %Sequences & ($3 \alpha$ fold)        & ($4 \beta + \alpha$ fold) \\ \hline
%GA95       & $0.000$                  & $2.901 \pm 0.475$         \\
%GB95       & $4.365 \pm 0.460$        & $0.000$                   \\
%GA88       & $0.000$                  & $ 3.941 \pm 0.517$        \\
%GB88       & $5.576 \pm 0.464$        & $0.000$                   \\
%GA98       &  &                           \\ \hline
%GB98       &                          &                           \\ \hline
%\end{tabular}
%\end{center}
%\end{table}

@@

This method is particularly well suited to pick out the native/native like structure from misfolded or decoys. We have
tested this using different models from CASP (critical assessment of structure
prediction). CASP is a blind test in which different groups world wide apply their
methods to predict the 3D structure of proteins given their sequence. This is done with a strict 3
week limit on each target (3 day for servers) and each group is allowed to submit 5 possible
structures (ranked from best to worst). In our role as assessors during the CASP8 and CASP9 \cite{MacCallum2011}
competition we observed no correlation at all between the ordering of structures submited by the
groups and the real ranking compared to an experimental model \cite{Kryshtafovych2011}. The
consequences of this goes beyond
those five structures; the deeper meaning is that groups producing ensembles of structures are
generating structures that are better than the ones they submit, but they do not know about it. In
fact, when querrying different groups after the results were known, most groups agree that this is
the case. Beyond the CASP problem, this reflects on the modelers ability to correctly rank order
models in many different environments, such as structures for drug design leads or designing more
stable proteins or peptide mimetics such as peptoids. One of the main culprits of this lack of
accuracy is the fact that ranking is often done via a potential energy function, which in many cases
lacks an entropy component. Other initiatives including knowledge based potentials do use some sort
of free energy to rank order, but its accuracy is not enough. Our method provides a physics based
solution to this problem. We have tried two experiments centered on CASP. First we tried to rank
order some structures from the previous CASP9 experiment. We have tested different options for that purpose.
First, we rank order submitions from same group and then from different groups. Our second test was 
to check whether this method can do better quality assessment/metaprediction than the best group in that catagory in CASP9. 
In both methods, our ranking is determined as a free energy measure,
and it is compared with the ranking given by a geometrical comparison between native and submitted
models called GDT\_TS (Global distance test score \cite{Zemla2003}). 
GDT\_TS represents the average percentage of residues that are in close proximity in two
structures optimally superimposed using four different distance cutoffs (1, 2, 4 and 8 Å).

We have chosen the inital models on which to test the methods based on server groups that have
traditionally done good in past CASP events. There is a high correlation between the structures that
are chosen for this method and its predictive accuracy. In particular, when GDT\_TS scores are below
50, it is difficult for us to say anything about the models. Our approach is simple, for every
subset of structures we choose, we use the confinement method in a pairwise procedure, and then rank
the structures. In all cases, we have also calculated the free energy with respect to the native
structure to check the stability of the models.
However, this step is not needed if we just want to rank order.

\subsubsection{The confinement method can rank different models generated by the same methodology and distinguish the native structure}

The question that we want to answer here is whether we can rank models that have been generated with
the same methodology. In all cases, we have calculated the free energy
difference between the native structure and the submitted models as well. 
%Identifying the native state from a set of models is a very easy test, that most metapredictors
%already handle correctly, but it is also a first proof of principle. 
First we choose a protein BVU3908 from Bacteroides vulgatus whose PDB id and CASP target code are
2L01 and T0559 respectively. The native structure of this 69 residue protein was solved using NMR.
The best predictor group for this target was "BAKER-ROSETTASERVER".
We initially checked similarity between the models submitted by "BAKER-ROSETTASERVER" and discarded two of them from the 
analysis on the basis of being too similar to some of the other models. The GDT\_TS score and rmsd
values as shown in Figure~\ref{fig:T0559} indicate that model 1 was predicted correctly, 
whereas the order of model 3 and 5 was wrong. The main difference
between model 3 and the rest of the models is that the orientation of the first alpha helix of model
3 is \Alberto{opposite??}.
On the other hand, as shown in Figure~\ref{fig:T0559} the confine and release method not only can differentiate the native 
structure from the submitted models, the ranking also correlates perfectly with the GDT\_TS score.

To further test the method we have taken the example of protein BT2368 from Bacteroides thetaiotaomicron. The PDB id of this 
74 amino acid residue protein is 2L02 and CASP target code is T0560. We have compared the free energy difference between 
the native structure and the two of the five submitted models from the group "Splicer". The remaining three models are discarded as 
again they are similar to the rest of the models. The comparison of GDT\_TS score as shown in 
Figure~\ref{fig:T0560} and published in the final CASP result is only for models with residues 3 to 66. In order to keep
consistency, we have also done our analysis with models and crystal structure consisting residues 3 to 66. As expected,
the native state is identified correctly, and the two other models are ranked in the correct order in accordance with the GDT\_TS score.

\subsubsection{The confinement method can rank models generated by different methods}

Our next test was between models that are produced with different methodologies. We do this by
looking at some of the best CASP submissions from different research groups. 
For this purpose we have chosen the x-ray crystallographic structure of 
fas apoptosis inhibitory protein molecule whose pdb id and CASP target codes are 3MX7 and T0540 respectively.
This protein contain 90 amino acid residues with 8 beta strands. We have chosen best models from groups
"LTB" and "Mufold" for our analysis, which are labelled as Model 1 and Model 2 in Figure~\ref{fig:T0540}.
As presented in  Figure~\ref{fig:T0540} we found that this method is able to correctly order the models which matches
with the GDT\_TS score.

\subsubsection{Per residue free energy calculation identifies the residues that are responsible for
differences between two conformations}

We used one of the CASP targets (pdb id=2KWY, casp id=T0569) to study the per residue contribution
between this NMR structure and the best CASP9 predicted model (from the 'Mufold' group, GDT\_TS=78).
Visualization upon superimposition of two structures reveals that
the difference between the two structure is at the region consist of residues 48 to 65, where model1 is much more disordered compare to the
native structure(see Figure~\ref{fig:T0569}). Other regions look similar for both the cases with of model1 is within 2.6 \AA backbone
rmsd value of the native structure. We found that, the confinement method can still differentiate between the native structure and 
the generated model. We have also calculated the per residue free energy with a aim that this can identify the residue which are responsible
for such free enerfy difference. The per residue free energy decomposition is shown in  Figure~\ref{fig:T0569_per_residue}(A). We indeed 
found that two hydrophobic residues Val-59 and Ile-61 are oriented towards the solvent in the protein 
surface in the generated model (Figure~\ref{fig:T0569_per_residue}(B). This the region, where there is a beta sheet in the original crystal structure, which is absent in the 
model. Interestinlgy, some regions are more stable in the generated model. A close inspection
reveals that 
the hydrophilic Glu in position 67 is completely exposed to the protein surface and solvent in the best model, whereas it is partially exposed in the 
crystal (Figure~\ref{fig:T0569_per_residue}(C)). The peak at position 76 is due to the H-bond between Lys-76 and Asp-11, 
which is absent in the original crystalographic structure (Figure~\ref{fig:T0569_per_residue}(D)).

\subsubsection{Failures in the confinement method}
